# Interpreting random forest

Since there is no inherent structure in a random forest, it is not immediately explainable. We are, however, interested in getting some idea about the overall impact of different features on the claim amount. For this reason we focus on making partial dependence plots.


## Partial dependence plots
We make pdps for continuous variables. This is not really sensible for unordered variables such as Gender and Maristat, and categorical variables like max speed (since this is defined by intervals).

```{r,include=FALSE}
loadData()
train_new <- train[,-c(3,4)]
train_new$ClaimInd<-as.factor(train_new$ClaimInd)

seq_ran<-readRDS("sequential_ranger_trained")
seq_ran_explain<-explain_mlr3(seq_ran,data=train_new[,-17],y=train_new[,17])
saveRDS(seq_ran_explain,"sequential_ranger_explainer")
seq_ran_explain<-readRDS("sequential_ranger_explainer")

seq_ran_pdp<-plot(model_profile(seq_ran_explain,variables=c("Exposure","BonusMalus","DrivAge","HasKmLimit","LicAge","RiskVar"),type="partial"))
seq_ran_pdp+ggtitle("Partial dependence plot for ranger",subtitle="")
```

## Histograms for categoricals

<!-- Denne kode er enormt grim -->
```{r,include=FALSE}
train_new$pred_claim_amount<-readRDS("sequential_ranger_prediction")$response

mean_predictions <- train_new %>%
  group_by(Gender) %>%
  summarise(MeanPrediction = mean(pred_claim_amount, na.rm = TRUE))

# Step 2: Calculate standard error and confidence intervals manually
summary_df <- train_new %>%
  group_by(Gender) %>%
  do({
    data_group <- .
    n_size <- nrow(data_group)
    mean_pred <- mean(data_group$pred_claim_amount, na.rm = TRUE)
    std_error <- sd(data_group$pred_claim_amount, na.rm = TRUE) / sqrt(n_size)
    lower_ci <- mean_pred - (std_error * 1.96)
    upper_ci <- mean_pred + (std_error * 1.96)
    
    data.frame(MeanPrediction = mean_pred, StdError = std_error, Lower = lower_ci, Upper = upper_ci)
  })
library(ggplot2)

ggplot(summary_df) +
  # Points with increased size and color differentiation
  geom_point(mapping=aes(x=Gender, y=MeanPrediction, color=Gender), size=4, alpha=0.8) +
  
  # Error bars with adjusted width and color
  geom_errorbar(mapping=aes(x=Gender, y=MeanPrediction, ymin=Lower, ymax=Upper), 
                width=0.2, color="darkgray", size=0.8) +
  
  # Customizing the plot with labels and title
  labs(x = "Gender", 
       y = "Mean Prediction", 
       title = "Mean claim amount by Gender",
       subtitle = "Includes 95% confidence intervals") +
  
  # Theme customization for a cleaner look
  theme_minimal(base_size = 14) +
  theme(legend.position = "none",  # Remove legend if color differentiation by Gender is not needed
        plot.title = element_text(face = "bold", size = 16),
        plot.subtitle = element_text(size = 12),
        axis.title.x = element_text(vjust = -0.2),
        axis.title.y = element_text(vjust = 1.2)) +
  
  # Optional: Customize colors manually if desired
  scale_color_manual(values = c("Male" = "#1f77b4", "Female" = "#ff7f0e"))

```

```{r,include=FALSE}
mean_predictions <- train_new %>%
  group_by(MariStat) %>%
  summarise(MeanPrediction = mean(pred_claim_amount, na.rm = TRUE))

# Step 2: Calculate standard error and confidence intervals manually
summary_df <- train_new %>%
  group_by(MariStat) %>%
  do({
    data_group <- .
    n_size <- nrow(data_group)
    mean_pred <- mean(data_group$pred_claim_amount, na.rm = TRUE)
    std_error <- sd(data_group$pred_claim_amount, na.rm = TRUE) / sqrt(n_size)
    lower_ci <- mean_pred - (std_error * 1.96)
    upper_ci <- mean_pred + (std_error * 1.96)
    
    data.frame(MeanPrediction = mean_pred, StdError = std_error, Lower = lower_ci, Upper = upper_ci)
  })
library(ggplot2)

ggplot(summary_df) +
  # Points with increased size and color differentiation
  geom_point(mapping=aes(x=MariStat, y=MeanPrediction, color=MariStat), size=4, alpha=0.8) +
  
  # Error bars with adjusted width and color
  geom_errorbar(mapping=aes(x=MariStat, y=MeanPrediction, ymin=Lower, ymax=Upper), 
                width=0.2, color="darkgray", size=0.8) +
  
  # Customizing the plot with labels and title
  labs(x = "Mariage Status", 
       y = "Mean Prediction", 
       title = "Mean claim amount by Mariage Status",
       subtitle = "Includes 95% confidence intervals") +
  
  # Theme customization for a cleaner look
  theme_minimal(base_size = 14) +
  theme(legend.position = "none",  # Remove legend if color differentiation by Gender is not needed
        plot.title = element_text(face = "bold", size = 16),
        plot.subtitle = element_text(size = 12),
        axis.title.x = element_text(vjust = -0.2),
        axis.title.y = element_text(vjust = 1.2)) +
  
  # Optional: Customize colors manually if desired
  scale_color_manual(values = c("Alone" = "#1f77b4", "Other" = "#ff7f0e"))
```


```{r,include=FALSE}
mean_predictions <- train_new %>%
  group_by(interaction(MariStat,Gender)) %>%
  summarise(MeanPrediction = mean(pred_claim_amount, na.rm = TRUE))

# Step 2: Calculate standard error and confidence intervals manually
summary_df <- train_new %>%
  group_by(interaction(MariStat,Gender)) %>%
  do({
    data_group <- .
    n_size <- nrow(data_group)
    mean_pred <- mean(data_group$pred_claim_amount, na.rm = TRUE)
    std_error <- sd(data_group$pred_claim_amount, na.rm = TRUE) / sqrt(n_size)
    lower_ci <- mean_pred - (std_error * 1.96)
    upper_ci <- mean_pred + (std_error * 1.96)
    
    data.frame(MeanPrediction = mean_pred, StdError = std_error, Lower = lower_ci, Upper = upper_ci)
  })
library(ggplot2)

summary_df$`interaction(MariStat, Gender)`

ggplot(summary_df) +
  # Points with increased size and color differentiation
  geom_point(mapping=aes(x=`interaction(MariStat, Gender)`, y=MeanPrediction, color=`interaction(MariStat, Gender)`), size=4, alpha=0.8) +
  
  # Error bars with adjusted width and color
  geom_errorbar(mapping=aes(x=`interaction(MariStat, Gender)`, y=MeanPrediction, ymin=Lower, ymax=Upper), 
                width=0.2, color="darkgray", size=0.8) +
  
  # Customizing the plot with labels and title
  labs(x = "Marriage Status and Genders", 
       y = "Mean Prediction", 
       title = "Mean claim amount by MariStatus.Gender",
       subtitle = "Includes 95% confidence intervals") +
  
  # Theme customization for a cleaner look
  theme_minimal(base_size = 14) +
  theme(legend.position = "none",  # Remove legend if color differentiation by Gender is not needed
        plot.title = element_text(face = "bold", size = 16),
        plot.subtitle = element_text(size = 12),
        axis.title.x = element_text(vjust = -0.2),
        axis.title.y = element_text(vjust = 1.2))+
  
  # Optional: Customize colors manually if desired
  scale_color_manual(values = c("Alone.Female" = "darkred", "Other.Female" = "salmon",Alone.Male="darkblue",Other.Male="cyan"))
```




